Verifying Concurrent Search Structure Templates for C Programs using VST

### 2023-04-03 concurrency/CONCUR 2023/paper.tex
working on section SS templates

\section{Introduction}
\label{sec:introduction}
Krishna et al. proposed concurrent search structure templates~\cite{templates} as a method for separating the proof of correctness of a concurrent access method (optimistic concurrency, hand-over-hand locking, and internal links) from the proof of correctness of the underlying data structure (linked list, hashtable, and B-tree). Ideally, it should be possible to prove the correctness of $n$ (single-threaded) data structure implementations and $m$ concurrency patterns and obtain $n * m$ verified concurrent data structures. In practice, the story is more complicated: certain patterns only work for specific data structures or require the data structures to store extra information, while some internal data structure operations may not fit the template model. In this paper, we apply the template approach to concurrent data structure implementations in C, verified using the Verified Software Toolchain (VST)~\cite{plfcc}, and report on its effectiveness and the challenges we encountered. The template approach depends crucially on the idea of \emph{logically atomic specifications} introduced in TaDA~\cite{tada} and further developed in Iris~\cite{iris}; our proofs take advantage of recent work integrating Iris-style logical atomicity into VST~\cite{mansky2022bringing}.

\section{Background}
\subsection{Concurrent Search Structure Templates}

@@

\section{Search Structure Templates}
\subsection{Lock Coupling}
\begin{figure}[!ht]
	\begin{subfigure}[t]{0.48\textwidth}
		\lstinputlisting[language=C, style=myStyle]{lock_traverse.c} 
	\end{subfigure}\qquad
		\begin{subfigure}[t]{0.49\textwidth}
			\lstinputlisting[language=C, style=myStyle]{lock_insert.c} 
	\end{subfigure}
	\caption{Code outlines for \lstinline{traverse} and \lstinline{insert}}
	\label{bst-conc}
\end{figure}
\newpage
\subsection{Give-Up}

The give-up template does not maintain locks during traversal between nodes, and instead only acquire them when loading or storing to a node. The range field, which is stored in each node, serves as a reference point for threads to determine if they are still on the correct path. If a thread finds that it has deviated from the expected path, it abandons the node by releasing the lock and returns to the root node to start again. This behavior is in line with the give-up template algorithm's approach, where threads terminate their search if they encounter obstacles or deviate from the expected path.

>>>>

\section{Introduction}
\label{sec:introduction}
Krishna et al. proposed concurrent search structure templates~\cite{templates} as a method for separating the proof of correctness of a concurrent access method (optimistic concurrency, hand-over-hand locking, and internal links) from the proof of correctness of the underlying data structure (linked list, hashtable, and B-tree). Ideally, it should be possible to prove the correctness of $n$ (single-threaded) data structure implementations and $m$ concurrency patterns and obtain $n * m$ verified concurrent data structures. However, in practice, the story is more complicated: certain patterns work only for specific data structures or require the data structures to store extra information, while some internal data structure operations may not fit the template model. In this paper, we apply the template approach to concurrent data structure implementations in C, which we verify using the Verified Software Toolchain (VST)~\cite{plfcc}, and report on its effectiveness and the challenges we encountered. The template approach depends crucially on the idea of \emph{logically atomic specifications} introduced in TaDA~\cite{tada} and further developed in Iris~\cite{iris}; our proofs take advantage of recent work integrating Iris-style logical atomicity into VST~\cite{mansky2022bringing}.

\section{Background}
\subsection{Concurrent Search Structure Templates}
A search structure is a type of data structure that is designed to facilitate the efficient storage and retrieval of data based on specific search criteria. These structures organize data in a manner that enables quick and efficient search, insertion, deletion, and traversal operations. Concurrent search structures, in turn, are data structures that can be accessed and modified concurrently by multiple threads or processes. The design of concurrent search structures poses several significant challenges, including the need to ensure correctness and consistency under concurrent access, achieve scalability by minimizing contention and maximizing parallelism, maintain performance and efficiency while managing synchronization and memory, and ensure compatibility with diverse hardware and software platforms. The complex interplay between concurrency and shared memory in concurrent search structures makes formal verification a challenging task. Modularity is a crucial aspect of simplifying formal proofs for concurrent search structures, as it is for designing and maintaining large systems. By decomposing complex systems into smaller, more manageable components, modularity facilitates the separation of concerns and minimizes interactions between different parts of the system, thus reducing the complexity of formal proofs. Moreover, modularity permits the reuse of verified components, which further simplifies the verification process.
{\color{red}{(Say something to link to template algorithms)}}

@@

\section{Search Structure Templates}
The C implementation of a node in our template style is:
\begin{lstlisting}[language = C, backgroundcolor=\color{white},   
	commentstyle=\color{codegreen},
	basicstyle=\ttfamily\footnotesize,
	breakatwhitespace=false,  breaklines=true,  keepspaces=true,                                
	showspaces=false,  showstringspaces=false,
	showtabs=false, tabsize=1]
typedef struct tree {int key; void *value; struct tree_t *left, *right;} tree;
typedef struct tree_t {tree *t; lock_t *lock; int min; int max;} tree_t;
typedef struct pn {struct tree_t *p; struct tree_t *n;} pn;
\end{lstlisting}
Each node (of type \lstinline{tree_t}) contains a \lstinline{lock} field that protects the node and a \lstinline{t} field. The \lstinline{t} field can be either \lstinline{NULL}, indicating that the node is a leaf node with no key or value, or it can point to a \lstinline{tree} struct that stores the node's key, value, and child pointers.

\subsection{Lock Coupling}
The lock-coupling template employs the hand-over-hand locking scheme to prevent interference from other threads during traversal. Each thread continuously maintains at least one lock during traversal between nodes, ensuring that other threads cannot interfere with the traversal or invalidate the ongoing search. Figure \ref{traverse_lock} presents two versions of the \lstinline{traverse} function: one written in C (Figure \ref{traverse_lock_a}) and the other in an ML-like language (Figure \ref{traverse_lock_b}).

\begin{figure}[!ht]
	\begin{subfigure}[t]{0.48\textwidth}
		\lstinputlisting[language=C, style=myStyle]{lock_traverse.c} 
		\caption{The \lstinline{traverse} method of the lock-coupling template give-up template algorithm written in C}
		\label{traverse_lock_a}
	\end{subfigure}\qquad
	\begin{subfigure}[t]{0.45\textwidth}
		\lstinputlisting[language=caml, style=myStyle]{lock_traverse.ml} 
		\caption{The \lstinline{traverse} method of the lock-coupling template algorithm written in ML-like language}
		\label{traverse_lock_b}	
	\end{subfigure}
	\caption{The \lstinline{traverse} method of the lock-coupling template algorithms}
	\label{traverse_lock}
\end{figure}

\begin{figure}[!ht]
	\begin{subfigure}[t]{0.48\textwidth}
		\lstinputlisting[language=C, style=myStyle]{giveup_traverse.c} 
		\caption{The \lstinline{traverse} method of the give-up template algorithm written in C}
	\end{subfigure}\qquad
	\begin{subfigure}[t]{0.45\textwidth}
		\lstinputlisting[language=caml, style=myStyle]{giveup_traverse.ml} 
		\caption{The \lstinline{traverse} method of the lock-coupling template algorithm}	
	\end{subfigure}
	\caption{The \lstinline{traverse} method of the give-up template algorithms}
	\label{traverse}
\end{figure}

\subsection{Give-Up}

The give-up template does not maintain locks during traversal between nodes; instead, it acquires them only when loading or storing to a node. In each node, the range field serves as a reference point for threads to determine if they are still on the correct path. If a thread finds that it has deviated from the expected path, it abandons the node by releasing the lock and returns to the root node to start again. This behavior aligns with the give-up template algorithm's approach, wherein threads terminate their search upon encountering obstacles or deviating from the expected path.
